Uppangala tree functionnal traits: Preliminray Results

### 2016-06-23 inst/doc/PreliminaryResults.Rmd
v0.1.9013 FD, all null models and result tables, and tree heights

# *News*
Add from the june 20th version :

- RLQ to investigate individuals filters
- Spatial autocorrelation for CWM and models residuals with Moran's I
- Phylosignal
- CWV
- Structural Equation Modelling (SEM)

@@

Figure 1: Principal component analysis (PCA) performed on individual trait means. *Left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line). See table 1 for abbreviation.*

@@

We can see that as wood density ans specific leaf area are orthogonal following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$). Leaf area is almost opposed to wood density and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007. This relationship could be explained by an equilibrium in tree hydraulics (Coomes et al, 2008). Finally, we found the positive relation suggested by @diaz_global_2015 between leaf dry matter content and wood density (Spearman's $rho = 0.24$, $p-value < 0.001$).

@@

```{r 4 env-com}
library(raster)
Env <- genEnv()
Env$Rivers <- readAll(Env$Rivers)
Trees <- genTrees()
Species <- genSpecies()
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
Trees$SpCode <- tolower(Trees$SpCode)
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
XY <- XY[-which(is.na(XY$X)),] # Removing NA
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
com@data$id <- paste0('C', com@data$id)
com@data$DEM <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'DEM'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'aspect'))
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$SouthWesterness <- sin((com@data$Aspect + 225) / 180 * pi)
com@data$NorthEasterness <- cos((com@data$Aspect + 225) / 180 * pi)
com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$topo <- as.factor(with(com@data, topo_classes(Slope, Curvature, Aspect)))
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
rm(Deriv, XY)
```

To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (aspect in direction of south-west), and distance to the nearest potential river. Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$. Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect in degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008]. Distance to the nearest potential river was caluclated under ArcGis (cite).

In order to avoid multi-colinearity in future regressions, we reduced our explanatory variables via a principal component analysis, see figure 2. Curvature, profile curvature, plan curvature, and distance to potential rivers covary together and with the principal component analysis first axis (PCA1). Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables. As a conclusion, we choose to reduce our variable to 4: curvature to represent the water flow paths and soil stabilization areas, wetness to highlight water accumulation areas, southwesterness to represent monsoon winds [@bunyan_effect_2015] and slope.

```{r PCA Env}
pca.env <- princomp(~ DEM + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SouthWesterness, data = com@data, cor = T)
lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
pca.env.var = apply(pca.env$score, 2, var)
plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
Hmisc::subplot(function(){
  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
          cex.names = 0.7, axes = T, names.arg = NA) ;
  lines(b, bstick(length(pca.env.var)), col = 'red') ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.05,0.35), from='npc'),
  y=grconvertY(c(0.025,0.30), from='npc')
)
par(new=T)
Rot <- t(t(pca.env$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, lwd = 2)
lines(c(0,0), XLIM, lty = 2, lwd = 2)
col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
        Rot[,1],Rot[,2],lwd=2,col=col)
legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
       text.col = col, cex = 1.2)
text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 1), rownames(Rot),cex=1.2,col=col)
axis (3)
axis (4)
```

Figure 2: Principal component analysis (PCA) performed on topgraphical variables. *Left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line).*

@@

To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density (StD) *(we may add tree height and growth in the future)*. Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$. Both basal area and stem density are calculated with 2013 Uppangala plots girth inventory only on stem with a diameter above 10 centimeters at breast height ($DBH>10$). Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$). So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).

@@

```{r 5 Com}
com <- CWT(Trees, com)
```

Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats. For all those community, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes representing the leaf and the wood economic spectrum (WES, LES). Weighted mean and variances were realized by abundance, presence-absence, or basal area.

Spatial variograms of community weighted means and variances didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$.

@@

```{r SEM}
library(lavaan)
m1 <- '
Abiotic =~ Slope+Curvature+Wetness+SouthWesterness
Traits =~ WD.cwm+Thick.cwm+LA.cwm+LDMC.cwm+SLA.cwm
Biotic =~ BA+Canopy
Traits ~ Abiotic+Biotic
'
m2 <- '
Abiotic =~ Slope+Curvature+Wetness+SouthWesterness
Traits =~ WD.cwm+Thick.cwm+LA.cwm+LDMC.cwm+SLA.cwm
Biotic =~ BA+Canopy
Biotic ~ Traits +Abiotic
'
fit1 <- sem(m1, data = com$abundance@data)
fit2 <- sem(m2, data = com$abundance@data)
semPlot::semPaths(fit1,what='path',whatLabels='std',intercept=FALSE,residuals=FALSE,thresholds=FALSE,nCharNodes=0,nCharEdges=0,groups='manlat',pastel=TRUE,title=TRUE,curveAdjacent=TRUE,ask=FALSE)
```

Figure 5: Path diagram for structural equation modelling. *Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic*.

Links between both community weighted means and variances with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables. Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables (slope, curvature, wetness and southwesterness for abiotic factors, and canopy height and basal area for biotic one). A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix). Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance. Results for abundance, presence-absence and basal area community weighted mean for all traits are presented in tables 4 and 5 (respectively for community weighted means and variances).

An assessment of spatial autocorrelation (SAC) of the community weighted means was conducted using Moran's I statistic. Most of community weighed means, especially in abundance, were significantly positively spatially autocorrelated (see table 3). But half of presence-absence and basal area community weighted means spatial autocrorrelation didn't significantly differ from 0 and the other were all low (Moran's I from 0.01 to 0.03). Specific leaf area and wood density are slightly spatially correlated suggesting a spatial filter due to environment or local history pressure (as dispersion limitation). Leaf dry mater content and more generally the wood economic spectrum possess a stochastic spatial distribution implying that they're not shaped by factors with a spatial organization. Finally the leaf economic spectrum, thickness and area are spatially organized for abundant species and voluminous individual but not in presence absence, suggesting that the filter is linked to abundant species pressure as dispersion filter.

```{r SAC}
library(ape)
traits <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
traits <- c(paste0(traits, '.cwm'), paste0(traits, '.cwv'))
weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
diag(weights) <- 0
SAC <- lapply(com, function(x){lapply(as.list(x@data[traits]), Moran.I, weights)})
SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
SAC <- data.frame(lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))}))
names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
SAC$Trait <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA')
SAC$Type <- ''
SAC$Type[c(1,8)] <- c('CWM', 'CWV')
SAC <- SAC[c('Type', 'Trait', 'Abundance', 'Presence-absence', 'Basal area')]
kable(SAC, row.names = F)
```

@@

```{r 11 lm & null models}
reg.abiot <- '~ Slope + Curvature + Wetness + SouthWesterness'
reg.biot <- '~ Canopy + BA'
# LM to simplify with loop or lapply
LM <- lapply(com, function(x) {with(x@data, list(LES.abiot.cwm = lm(formula(paste('LES.cwm', reg.abiot))), # Abiotic CWM
                                                 logSLA.abiot.cwm = lm(formula(paste('log(SLA.cwm)', reg.abiot))), 
                                                 WES.abiot.cwm = lm(formula(paste('WES.cwm', reg.abiot))),
                                                 WD.abiot.cwm = lm(formula(paste('WD.cwm', reg.abiot))),
                                                 LDMC.abiot.cwm = lm(formula(paste('LDMC.cwm', reg.abiot))),
                                                 logLA.abiot.cwm = lm(formula(paste('log(LA.cwm)', reg.abiot))),
                                                 Thick.abiot.cwm = lm(formula(paste('Thick.cwm', reg.abiot))),
                                                 LES.abiot.cwv = lm(formula(paste('LES.cwv', reg.abiot))), # Abiotic CWV
                                                 logSLA.abiot.cwv = lm(formula(paste('log(SLA.cwv)', reg.abiot))), 
                                                 WES.abiot.cwv = lm(formula(paste('WES.cwv', reg.abiot))),
                                                 WD.abiot.cwv = lm(formula(paste('WD.cwv', reg.abiot))),
                                                 LDMC.abiot.cwv = lm(formula(paste('LDMC.cwv', reg.abiot))),
                                                 logLA.abiot.cwv = lm(formula(paste('log(LA.cwv)', reg.abiot))),
                                                 Thick.abiot.cwv = lm(formula(paste('Thick.cwv', reg.abiot))),
                                                 LES.biot.cwm = lm(formula(paste('LES.cwm', reg.biot))), # Biotic CWM
                                                 logSLA.biot.cwm = lm(formula(paste('log(SLA.cwm)', reg.biot))),
                                                 WES.biot.cwm = lm(formula(paste('WES.cwm', reg.biot))),
                                                 WD.biot.cwm = lm(formula(paste('WD.cwm', reg.biot))),
                                                 LDMC.biot.cwm = lm(formula(paste('LDMC.cwm', reg.biot))),
                                                 logLA.biot.cwm = lm(formula(paste('log(LA.cwm)', reg.biot))),
                                                 Thick.biot.cwm = lm(formula(paste('Thick.cwm', reg.biot)))
                                                 ))})
LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
model <- null_model(com, Trees, PFT_sp, lapply(LM, function(y){lapply(y, function(x){x$call})}), SpCode, n = 999, verbose = F, time = T)
```

## Abiotic filters

```{r CWM abiotic results}
kable(regression_table(LM, model, reg.abiot, select = '.abiot.cwm'))
```

Table 4: Step mesure on abiotic factors and null models summary for traits community weighted means. *Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW). Model deviance was compared to 999 null models deviance in which traits were randomized among species.*

@@

Leaf economic spectrum (LES) and specific leaf area (SLA) is associated non-randomly with convex and dry surfaces, as the thickness for basal area community weighted mean. Trees will tend to increase their surface compare to their carbon investment in areas where the water will flow but not accumulate nor stay, and where the soil will tend to accumulate and be stabilized. In favourable water and soil conditions, trees will tend to invest more in leaf ressource capture at the expense of leaf defense, adopting the acquisitive [@sterck_functional_2011] our fast growing strategy [@reich_world-wide_2014].

Wood density (WD) in presence absence shows a non-random association with southwest steep slope and concave surfaces with a relatively important explanation level ($R^2=0.23$). Trees will increase their wood density in thin soil and water and soil departure conditions under potential monsoon winds *(We need to decide or not if regarding the plot condition we can accept or not the southwest orientation as being linked to monsoon winds)*. In water stress and mechanical hazard conditions, trees will compensate by increasing their water transport safety and mechanical resistance and resilience with an increased wood density [@chave_towards_2009].

```{r SAC residuals}
library(ape)
weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
diag(weights) <- 0
SAC <- lapply(LM, function(x){lapply(x, function(y){Moran.I(residuals(y), weights)})})
SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
SAC <- data.frame(lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))}))
names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
row.names(SAC) <- row.names(model$p.value)
# kable(SAC)
```

Model residuals spatial autocorrelation (SAC) is null confirming that our models explain fully the spatial structur of our 5 traits and 2 strategy axes (compare to table 3). At the exception of thickness that still shows a significant spatial autocorrelation, suggesting that our model missed spatially structured factors in this case.

```{r CWV abiotic results}
kable(regression_table(LM, model, reg.abiot, select = '.abiot.cwv'))
```

>>>>

# *News*
Add from the `r Sys.Date()` version :

- RLQ to investigate individuals filters
- Spatial autocorrelation for CWM and models residuals with Moran's I
- Phylosignal
- CWV
- Structural Equation Modelling (SEM)
- Species growth too much unprecised for analysis
- Eveness and other FD package metrics

@@

Figure 1: Principal component analysis (PCA) performed on individual trait means. *Bottom right corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 6 components (red line). See table 1 for abbreviation.*

@@

We can see that as wood density and specific leaf area are orthogonal following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$). Leaf area is almost opposed to wood density and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007. This relationship could be explained by an equilibrium in tree hydraulics (Coomes et al, 2008). Finally, we found the positive relation suggested by @diaz_global_2015 between leaf dry matter content and wood density (Spearman's $rho = 0.24$, $p-value < 0.001$).

@@

```{r 4 env-com}
library(raster)
Env <- genEnv()
Env$Rivers <- readAll(Env$Rivers)
Trees <- genTrees()
Species <- genSpecies()
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
Trees$SpCode <- tolower(Trees$SpCode)
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
XY <- XY[-which(is.na(XY$X)),] # Removing NA
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
com@data$id <- paste0('C', com@data$id)
com@data$DEM <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'DEM'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'aspect'))
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$SW <- sin((com@data$Aspect + 225) / 180 * pi)
com@data$NE <- cos((com@data$Aspect + 225) / 180 * pi)
com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$topo <- as.factor(with(com@data, topo_classes(Slope, Curvature, Aspect)))
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
# com@data$Height <- aggregate(Trees$height, list(Trees$com), mean, na.rm = T)[,2]
rm(Deriv, XY)
```

To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river. Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$. Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect in degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008]. Distance to the nearest potential river was caluclated under ArcGis (cite).

In order to avoid multi-colinearity in future regressions, we reduced our explanatory variables via a principal component analysis, see figure 2. Curvature, profile curvature, plan curvature, and distance to potential rivers covary together and with the principal component analysis first axis (PCA1). Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables. As a conclusion, we choose to reduce our variable to 4: curvature to represent the water flow paths and soil stabilization areas, wetness to highlight water accumulation areas, southwesterness (SW) to represent monsoon winds [@bunyan_effect_2015] and slope.

```{r PCA Env}
pca.env <- princomp(~ DEM + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + Rivers + SW, data = com@data, cor = T)
lambda <- pca.env$sdev * sqrt(nrow(pca.env$scores))
pca.env.var = apply(pca.env$score, 2, var)
plot(t(t(pca.env$scores)/lambda),pch=16, col = 'lightgrey',
     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
     xlab = paste('Axe 1 -', round(pca.env.var[1]/sum(pca.env.var)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(pca.env.var[2]/sum(pca.env.var)*100, 2), '%'))
Hmisc::subplot(function(){
  b <- barplot((pca.env.var / sum(pca.env.var)), las = 2,
          cex.names = 0.7, axes = T, names.arg = NA) ;
  lines(b, bstick(length(pca.env.var)), col = 'red') ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.05,0.35), from='npc'),
  y=grconvertY(c(0.025,0.30), from='npc')
)
par(new=T)
Rot <- t(t(pca.env$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, lwd = 2)
lines(c(0,0), XLIM, lty = 2, lwd = 2)
col <- c("#1874CD", "#008B00", "#CD3700", "#9400D3")[c(2, 3, 1, 1, 1,  2, 1, 4)]
arrows (rep(0,nrow(pca.env$loadings)),rep(0,nrow(pca.env$loadings)),
        Rot[,1],Rot[,2],lwd=2,col=col)
legend('topright', c('Elevation', 'Slope', 'Water flow', 'South-West'),
       text.col = col, cex = 1.2)
text (Rot[,1:2], pos = c(2, 2, 2, 1, 1, 3, 3, 1), rownames(Rot),cex=1.2,col=col)
axis (3)
axis (4)
```

Figure 2: Principal component analysis (PCA) performed on topgraphical variables. *Bootom left corner subplot represents the eigen values (grey bars) compare to a broken stick distribution for 8 components (red line). SW stand for southwesterness.*

@@

To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density *(we may add tree height and growth in the future)*. Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$. Both basal area and stem density are calculated with 2013 Uppangala plots girth inventory only on stem with a diameter above 10 centimeters at breast height ($DBH>10$). Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$). So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).

@@

```{r 5 Com}
library(FD)
com <- CWT(Trees, com)
x <- data.frame(PFT_sp[-1], row.names = PFT_sp$Sp_Code)
a <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
a[is.na(a)] <- 0
x <- x[names(a),]
FD <- as.data.frame(dbFD(x, a, corr = 'cailliez', calc.CWM = F, print.pco = F, messages = F))[-c(1:2,4)]
names(FD) <- paste0(names(FD), '.fd')
FD$id <- row.names(FD)
com$abundance <- merge(com$abundance, FD)
```

Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats. For all those community, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes representing the leaf and the wood economic spectrum (WES, LES). Weighted mean and variances were realized by abundance, presence-absence, or basal area. In addition functionnal diversity indices weighted by abundances were computed with the `FD` package [@laliberte_distance-based_2010] for our communities: functional richness (FRic), functional evenness (FEve), and functional divergence (FDiv) and Rao's quadratic entropy (Q) (cite Villéger et al. (2008) et Botta-Dukát 2005).

Spatial variograms of community weighted means and variances didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$. Functionnal diversity indices show also a plateau around $70 m$.

@@

```{r SEM}
library(lavaan)
m1 <- '
Abiotic =~ Slope + Curvature + Wetness + SW
Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
Biotic =~ BA + Canopy
Traits ~ Abiotic + Biotic
'
m2 <- '
Abiotic =~ Slope + Curvature + Wetness + SW
Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
Biotic =~ BA + Canopy
Biotic ~ Traits +Abiotic
'
fit1 <- sem(m1, data = com$abundance@data)
fit2 <- sem(m2, data = com$abundance@data)
semPlot::semPaths(fit1,what='path',whatLabels='std',intercept=FALSE,residuals=FALSE,thresholds=FALSE,nCharNodes=0,nCharEdges=0,groups='manlat',pastel=TRUE,title=TRUE,curveAdjacent=TRUE,ask=FALSE)
```

Figure 5: Path diagram for structural equation modelling. *Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic. Arrow values are indicating standardized parameter estimates*.

Links between both community weighted means and variances with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables. Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables (slope, curvature, wetness and southwesterness for abiotic factors, and canopy height and basal area for biotic one). A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix). Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance. Results for abundance, presence-absence and basal area community weighted mean for all traits are presented in tables 4 and 5 (respectively for community weighted means and variances). Same procedure was also applied to functionnal diversity indices (FD) from our ocmmunities, results are presented in table 6.

An assessment of spatial autocorrelation (SAC) of the community weighted means and variances and functionnal diversity indices was conducted using Moran's I statistic. Only half of community weighted means and variances were significantly positively spatially autocorrelated (see table 3) with low values (Moran's I from 0.01 to 0.04). Leaf economic spectrum and dry matter content weighted means as wood density weighted variances are slightly spatially correlated suggesting a spatial filter due to environment or local history pressure (as dispersion limitation). On the opposite, wood density and thickness community weighted means as wood economic spectrum community weighted variances possess a stochastic spatial distribution implying that they're not shaped by factors with a spatial organization. Finally remaining traits community weighted means and variances are spatially organized for abundant species and voluminous individual but not in presence absence, suggesting that the filter is linked to abundant species pressure as dispersion filter. As a conclusion both community weighted means and variances are most of the time stochastic or driven by dispersion. Moreover all functionnal diversity indices (FD) are spatially correlated.

@@

```{r SAC}
library(ape)
traits <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
traits <- c(paste0(traits, '.cwm'), paste0(traits, '.cwv'))
traits <- list(c(traits, 'FRic.fd', 'FEve.fd', 'FDiv.fd', 'FDis.fd', 'RaoQ.fd'), traits, traits)
weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
diag(weights) <- 0
SAC <- mapply(function(x,y){lapply(as.list(x@data[y]), Moran.I, weights)}, x =com, y = traits)
SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
SAC <- lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))})
SAC[[2]] <- c(SAC[[2]], rep('', 5)) ; SAC[[3]] <- c(SAC[[3]], rep('', 5))
SAC <- data.frame(SAC)
names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
SAC$Indice <- c(rep(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA'), 2), c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ'))
SAC$Type <- ''
SAC$Type[c(1,8,15)] <- c('CWM', 'CWV','FD')
SAC <- SAC[c('Type', 'Indice', 'Abundance', 'Presence-absence', 'Basal area')]
kable(SAC, row.names = F)
```

```{r 6 lm & null models}
reg <- '~ Slope + Curvature + Wetness + SW + Canopy + BA'
# LM to simplify with loop or lapply
LM <- lapply(com, function(x) {with(x@data, list(LES.cwm = lm(formula(paste('LES.cwm', reg))), # CWM
                                                 logSLA.cwm = lm(formula(paste('log(SLA.cwm)', reg))), 
                                                 WES.cwm = lm(formula(paste('WES.cwm', reg))),
                                                 WD.cwm = lm(formula(paste('WD.cwm', reg))),
                                                 LDMC.cwm = lm(formula(paste('LDMC.cwm', reg))),
                                                 logLA.cwm = lm(formula(paste('log(LA.cwm)', reg))),
                                                 Thick.cwm = lm(formula(paste('Thick.cwm', reg))),
                                                 LES.cwv = lm(formula(paste('LES.cwv', reg))), # CWV
                                                 logSLA.cwv = lm(formula(paste('log(SLA.cwv)', reg))), 
                                                 WES.cwv = lm(formula(paste('WES.cwv', reg))),
                                                 WD.cwv = lm(formula(paste('WD.cwv', reg))),
                                                 LDMC.cwv = lm(formula(paste('LDMC.cwv', reg))),
                                                 logLA.cwv = lm(formula(paste('log(LA.cwv)', reg))),
                                                 Thick.cwv = lm(formula(paste('Thick.cwv', reg)))))})
LM$abundance <- c(LM$abundance, with(com$abundance@data, list(FRic.fd = lm(formula(paste('FRic.fd', reg))),
                                                              FEve.fd = lm(formula(paste('FEve.fd', reg))),
                                                              FDiv.fd = lm(formula(paste('FDiv.fd', reg))),
                                                              FDis.fd = lm(formula(paste('FDis.fd', reg))),
                                                              RaoQ.fd = lm(formula(paste('RaoQ.fd', reg))))))
LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
model <- null_model(com, Trees, PFT_sp, lapply(LM, function(y){lapply(y, function(x){x$call})}), SpCode, n = 999, verbose = F, time = T)
```

## Community weighted means

Table 4: Summary for linear models and null models for traits community weighted means. *Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA). Model deviance was compared to 999 null model deviances in which traits were randomized among species. See table 1 for abbreviations.*

```{r CWM}
kable(regression_table(LM, model, reg, select = '.cwm'))
```

### Abiotic factors

@@

Leaf economic spectrum (LES) and specific leaf area (SLA) is associated non-randomly with convex and dry surfaces, as the thickness for basal area community weighted mean. Trees will tend to increase their surface compare to their carbon invest ment in areas where the water will flow but not accumulate nor stay, and where the soil will tend to accumulate and be stabilized. In favourable water and soil conditions, trees will tend to invest more in leaf ressource capture at the expense of leaf defense, adopting the acquisitive [@sterck_functional_2011] our fast growing strategy [@reich_world-wide_2014].

Wood density (WD) in presence absence shows a non-random association with southwest steep slope and concave surfaces with a relatively important explanation level ($R^2=0.25$). Trees will increase their wood density in thin soil and water and soil departure conditions under potential monsoon winds *(We need to decide or not if regarding the plot condition we can accept or not the southwest orientation as being linked to monsoon winds)*. In water stress and mechanical hazard conditions, trees will compensate by increasing their water transport safety and mechanical resistance and resilience with an increased wood density [@chave_towards_2009].

### Biotic factors

Most of community weighted means don't show association with biotic factors, and most of the rare association are not significantly different from null models randomly generated. Still leaf area (LA) in basal area is associated with high trees. Voluminous individuals tend to increase their leaf surface in heigher canopy condition to compensate understory shadow. And wood density tends also to increase with tree height for rare species, suggesting that with higher canopy denser wood species will tend to grow in a community, as shade species.

## Community weighted variances

Table 5: Summary for linear models and null models for traits community weighted variances. *Linear model results for each trait community weighted variances (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA). Model deviance was compared to 999 null model deviances in which traits were randomized among species. See table 1 for abbreviations.*

```{r CWV}
kable(regression_table(LM, model, reg, select = '.cwv'))
```

## Functionnal diversity
